---
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls(all=TRUE))
knitr::opts_chunk$set(echo = TRUE, fig.path="figures/")
```

## Required packages

```{r}
library(deSolve)
```

# Ordinary differential equations

## Determinate inflorescence 

Write the ordinary differential equations for determinate inflorescences.

```{r}
controlDeterminate=function(t,y,parms,controlFunction) {
  Ve=y[1]; Le=y[2]; In=y[3]; Fl=y[4];  
  
  beta1=parms[1]; 
  beta2=parms[2];
  
  ut = controlFunction(t)
  
  derivs[1]= -beta1*ut*Ve ;
  derivs[2]= beta1*Ve;
  derivs[3]= beta1*ut*Ve-beta2*In;
  derivs[4]= beta2*In;

  return(list(derivs));
}
```

## Indeterminate inflorescence

Write the ordinary differential equations for indeterminate inflorescences.

```{r}
controlIndeterminate=function(t,y,parms,controlFunction) {
  Ve=y[1]; Le=y[2]; In=y[3]; Fl=y[4];  
  
  beta1=parms[1]; 
  beta2=parms[2];
  
  ut = controlFunction(t)
  
  derivs[1]= -beta1*ut*Ve ;
  derivs[2]= beta1*Ve;
  derivs[3]= beta1*ut*Ve;
  derivs[4]= beta2*In;

  return(list(derivs));
}
```

# Initial conditions

```{r }
# Vector to hold the derivatives
derivs=numeric(4); 

# starting values of both state variables
y0=c(Ve=1,Le=0,In=0,Fl=0); 
# time sequence
times=seq(0,8,length.out=200)
# controls
parms=c( beta1=1, 
         beta2=1);
```

# Function to plot solutions 

```{r}
plotSolutions = function(ode1=odeSolutionDeterminate){

par(mfrow=c(1,2))

plot(ode1$time,ode1$Ve,type="l",col="red",bty="n",
     #main="Meristem dynamics for annuals\n with a determinate inflorescence (solid)\n and indeterminate inflorescence (dotted).",
     xlab="Time (t)",
     ylab="Available meristems",
     ylim=c(0,1  ), 
     cex.lab=.5, cex.axis=.5, cex.main=.75, cex.sub=.5)

lines(ode1$time,ode1$In,col="blue")
#lines(ode2$time,ode2$In,col="blue",lty='dotted')
#lines(ode2$time,ode2$Ve,col="red",lty='dotted')

# legend(x = 0, y = max(ode1$In,ode2$In) , 
#        legend = c("Vegetative meristems (Ve)", "Inflorescence meristems (In)"), 
#        col = c('red', 'blue') ,
#        lty = c(1,1), 
#        cex = .25)

plot(ode1$time,ode1$Le,type="l",col="red",bty="n",
     #main="Growth and reproduction for annuals\n with a determinate inflorescence (solid)\n and indeterminate inflorescence (dotted).",
     xlab="Time (t)",
     ylab="Biomass ",
     ylim=c(0,1) , 
     cex.lab=.5, cex.axis=.5, cex.main=.75, cex.sub=.5)

lines(ode1$time,ode1$Fl,col="blue")

#lines(ode2$time,ode2$Fl,col="red",lty='dotted')
#lines(ode2$time,ode2$Le,col="blue",lty='dotted')

# legend(x = 0, y = max(ode1$Fl,ode2$Fl),
#        legend = c("Vegetative biomass (V)", "Reproductive biomass (F)"),
#        col = c('red', 'blue') ,
#        lty = c(1,1),
#        cex = .25)

#abline(h=1)
}

```

# Instantaneous switch

```{r instant-switch}
instantSwitch = function(t){
  if ( t <= 4 )
    tmp <- 0
  else
    tmp <- 1
  
  return(tmp)
}

odeSolutionDeterminate=ode(y0,times,controlDeterminate,parms,controlFunction=instantSwitch);
odeSolutionDeterminate = data.frame(odeSolutionDeterminate);

plotSolutions(odeSolutionDeterminate)

```

# Graded switch

```{r graded-switch}
gradedSwitch = function(t){
  if ( t <= 2)
    tmp <- 1
  else if (t <= 4)
    tmp <- .5
  else
    tmp <- 0
  
  return(tmp)
  }

odeSolutionDeterminate=ode(y0,times,controlDeterminate,parms,controlFunction=gradedSwitch);
odeSolutionDeterminate = data.frame(odeSolutionDeterminate);

gradedSwitch = function(t){
  
  tmp = c() # tmp[1] = pt, tmp[2] = qt
  
  if ( t <= 2)
    {tmp[1] <- 1
    tmp[2] <- 0}
  else if ( t <= 4)
  {tmp[1] <- .5
    tmp[2] <- 0}
  else 
  {tmp[1] <- 0
   tmp[2] <- 1}
  
  return(tmp)
  }

odeSolutionIndeterminate=ode(y0,times,controlIndeterminate,parms,controlFunction=gradedSwitch);
odeSolutionIndeterminate = data.frame(odeSolutionIndeterminate)

plotSolutions(odeSolutionDeterminate,odeSolutionIndeterminate)
```

# Graded switch

```{r}
gradedSwitch = function(t){
  if ( t <= 2)
    tmp <- .75
  else if (t <= 4)
    tmp <- .5
  else
    tmp <- 0
  
  return(tmp)
  }

odeSolutionDeterminate=ode(y0,times,controlDeterminate,parms,controlFunction=gradedSwitch);
odeSolutionDeterminate = data.frame(odeSolutionDeterminate);

gradedSwitch = function(t){
  
  tmp = c() # tmp[1] = pt, tmp[2] = qt
  
  if ( t <= 2)
    {tmp[1] <- .75
    tmp[2] <- 0}
  else if ( t <= 4)
  {tmp[1] <- .5
    tmp[2] <- 0}
  else 
  {tmp[1] <- 0
   tmp[2] <- 1}
  
  return(tmp)
  }

odeSolutionIndeterminate=ode(y0,times,controlIndeterminate,parms,controlFunction=gradedSwitch);
odeSolutionIndeterminate = data.frame(odeSolutionIndeterminate)

plotSolutions(odeSolutionDeterminate,odeSolutionIndeterminate)
```

## Instantaneous switch with resources

```{r}
instantSwitch = function(t){
  if ( t <= 4)
    tmp <- 1
  else
    tmp <- 0
  
  return(tmp)
  }

odeSolutionDeterminate=ode(y0,times,controlDeterminate,parms,controlFunction=instantSwitch);
odeSolutionDeterminate = data.frame(odeSolutionDeterminate);

odeSolutionResource=ode(y0,times,controlResource,parms,controlFunction=instantSwitch);
odeSolutionResource = data.frame(odeSolutionResource)

plotSolutions(odeSolutionDeterminate,odeSolutionResource)
```